---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# declusteringr

<!-- badges: start -->
<!-- badges: end -->

The goal of declusteringr is to declustering spatial data. This package offers support for `sf` spatial objects.

## Installation

You can install the released version of declusteringr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("declusteringr")
```

## Examples

This is a basic example of cell declustering:

```{r example}
library(declusteringr)
gridWeighted(spatial_object = sf::st_read('dane/punkty1.shp'), cellsize = 0.05)
```

`gridWeighted` is adding columns row.id - the cell's number, col.id - the point's number, received_weights - the weight of each cell and lengths - amounts of each points in cell. `gridWeighted` does not change amount of features.
Weights are calculated according to this pattern:



This is basic example of random declustering:

```{r}
gridRandom(spatial_object = sf::st_read('dane/punkty1.shp'), cellsize = 0.05 , numpoint = 1)
```

`gridRandom` is choosing random features by specifying some parameters. `gridRandom` can change amount of features.
